Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-12 pecomplexity.tex
added notion of redundancy to the abstract

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series—which re-
sults from the inherent dimension, nonlinearity, and non-stationarity
of the generating process, as well as from measurement issues like
noise, aggregation, and finite data length—is both empirically
quantifiable and directly correlated with predictability. To validate
this conjecture, we study 120 different time-series data sets and a
wide variety of forecast models. We estimate the complexity of each
data set using a variant of permutation entropy, then compare that
complexity with the accu- racy of each forecast method. We use the
results to develop a heuristic that can help practitioners recognize
when a particular prediction method is not well matched to the task at
hand: that is, when the time series has more predictive structure than
that method can capture and exploit.

>>>>

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series —-- which
results from the inherent dimension, nonlinearity, and
non-stationarity of the generating process, as well as from
measurement issues like noise, aggregation, and finite data length ---
is both empirically quantifiable and directly correlated with
predictability. To validate this conjecture, we study 120 different
time-series data sets.  We argue that \emph{redundancy} is an
effective way to measure complexity and predictive structure in a time
series and that \emph{weighted permutation entropy} is an effective
way to estimate that complexity.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a heuristic that can help
practitioners recognize when a particular prediction method is not
well matched to the task at hand: that is, when the time series has
more predictive structure than that method can capture and exploit.
